---
title: R Notebook HICP and Consumption share
---
```{r}
#install.packages('ecb')

#### Data from ECB ####
rm(list=ls())
library(ecb)
library(ggplot2)
library(dplyr)
library(reshape2)
```
LOAD DATA from ECB SDW
```{r}
names<-list("ICP.M.FR.N.000000.4.INX", "ICP.M.FR.N.010000.4.INX", "ICP.M.FR.N.020000.4.INX", "ICP.M.FR.N.030000.4.INX", "ICP.M.FR.N.040000.4.INX",	"ICP.M.FR.N.050000.4.INX", "ICP.M.FR.N.060000.4.INX", "ICP.M.FR.N.070000.4.INX",	"ICP.M.FR.N.080000.4.INX",	"ICP.M.FR.N.090000.4.INX",	"ICP.M.FR.N.100000.4.INX",	"ICP.M.FR.N.110000.4.INX",	"ICP.M.FR.N.120000.4.INX")

x<-get_data("ICP.M.FR.N.000000.4.INX",filter=list(startPeriod=2000,endPeriod=2021))
data=data.frame(as.Date(paste(x$obstime,1,sep="-"),"%Y-%m-%d"))

for (i in names){
  x<-get_data(i,filter=list(startPeriod=2000,endPeriod=2021))
  data<-cbind(data,x$obsvalue)
}

names(data)<-c('Date','HICP - Overall index',	'HICP - FOOD','HICP - ALCOHOL, TOBACCO','HICP - CLOTHING','HICP - ELECTRICITY, GAS','HICP - FURNISHINGS','HICP - HEALTH','HICP - TRANSPORT','HICP - COMMUNICATION','HICP - RECREATION	','HICP - EDUCATION	','HICP - RESTAURANTS','HICP - MISCELLANEOUS')

colnames(data) <- gsub('HICP - ', '', colnames(data), fixed=TRUE)
```

Get all categories price indices over time.

```{r}
d <- melt(data, id.vars="Date")
monthly<-d %>%
    group_by(variable) %>%
    mutate(VAL.IND = value/value[Date=="2000-01-01"])%>%
    mutate(Year=cut(Date,"year"))
  
log_month<- monthly %>%
  group_by(variable)%>%
  mutate(log_v=c(NA,diff(log(VAL.IND))))

annual<-monthly %>%
  group_by(Year, variable)%>%
  filter(row_number()==1)

log_annual<- annual %>%
  group_by(variable)%>%
  mutate(log_v=c(NA,diff(log(VAL.IND))))

pdf("PIndex_inflation.pdf")
ggplot(monthly, aes(x=Date,y=VAL.IND, col=variable)) + 
  geom_path()+
  theme_bw()+
  labs(x="Date",y="Monthly Index", title="Monthly price index with year 2000=1", caption = "ECB SDW Author's calculation, ECB SDW")

ggplot(log_month, aes(x=Date,y=log_v, col=variable)) + 
  geom_path()+
  theme_bw()+
  labs(x="Date",y="Monthly Inflation", title="Monthly Inflation for each categorie", caption = "ECB SDW Author's calculation, ECB SDW")

ggplot(annual, aes(x=Date,y=VAL.IND, col=variable)) + 
  geom_path()+
  theme_bw()+
  labs(x="Date",y="Annual Index", title="Annual price index with year 2000=1", caption = "ECB SDW Author's calculation, ECB SDW")

ggplot(log_annual, aes(x=Date,y=log_v, col=variable)) + 
  geom_path()+
  theme_bw()+
  labs(x="Date",y="Annual Inflation", title="Annual Inflation for each categorie", caption = "ECB SDW Author's calculation, ECB SDW")

dev.off()

```


```{r}
library(readxl)
#Average cons per purpose dataset
average_con <- read_excel("average_con.xlsx", 
    sheet = "Sheet 1")

colnames(average_con)<-c("Date","Type","Value")
average_con[,1]<-as.Date(average_con$Date, "%Y")
average_con[,3]<-as.numeric(average_con$Value)

#quintile dataset
cons_per_decile <- read_excel("cons_per_decile.xlsx", 
    sheet = "Sheet 1")

#Urbanisation dataset
cons_per_city <- read_excel("cons_per_city.xlsx", 
    sheet = "Sheet 1")

#CSP dataset
cons_per_csp <- read_excel("cons_per_csp_France.xlsx", 
    sheet = "Sheet 1")

#Age dataset
cons_per_age<-read_excel("cons_per_age.xlsx", 
    sheet = "Sheet 1")
```

Obtain consumption share for the average household.

```{r}
average_con<-average_con %>% filter(Type!="Total")
average_con<-average_con %>% filter(Date!="2021-09-03")

average_con$Type <- sapply(average_con$Type, switch, 
                      "Food and non-alcoholic beverages"="FOOD",
                      "Alcoholic beverages, tobacco and narcotics"="ALCOHOL, TOBACCO",
                      "Clothing and footwear"="CLOTHING",
                      "Housing, water, electricity, gas and other fuels"="ELECTRICITY, GAS",
                      "Furnishings, household equipment and routine household maintenance"="FURNISHINGS",
                      "Health"="HEALTH",
                      "Transport"="TRANSPORT",
                      "Communications"="COMMUNICATION",
                      "Recreation and culture" ="RECREATION",
                      "Education"="EDUCATION",
                      "Restaurants and hotels"="RESTAURANTS",
                      "Miscellaneous goods and services"="MISCELLANEOUS",
)
   

average_con<-average_con %>%
  group_by(Type, y=cut(Date, "year"))%>%
  mutate(mean_val=mean(Value))


png(file="cons_average.png")              
ggplot(average_con, aes(fill=Type, y=Value, x=Date)) + 
  geom_bar(position="stack", stat="identity")+
  theme_bw()+
  labs(y="Share", title = "Consumption share average household")+
  scale_fill_brewer(palette="Set3")
dev.off()
```

Prepare consumption per purpose by quintile dataset to obtain the overall consumption share by purpose and quintile.
```{r}
colnames(cons_per_decile)<-c("Type","income_grp","s_1999","s_2005","s_2010","s_2015")
d<-cons_per_decile %>%
    group_by(income_grp) %>%
    mutate(Somme_1999 = sum(s_1999), Somme_2005=sum(s_2005), Somme_2010=sum(s_2010), Somme_2015=sum(s_2015))

share<-d %>%
  group_by(income_grp) %>%
  summarize(Type=Type, s_1999/Somme_1999, s_2005/Somme_2005, s_2010/Somme_2010, s_2015/Somme_2015 )

colnames(share)<-c("income_grp","Type","s_1999","s_2005","s_2010","s_2015")
d <- melt(share, id.vars=c("Type","income_grp"))
d$variable <- gsub('s_', '', d$variable, fixed=TRUE)

colnames(d)<-c("Type", "income_grp", "Date", "Value")

d$Quintile <- sapply(d$income_grp, switch, 
                     "First quintile"="1",
                     "Second quintile"="2",
                     "Third quintile"="3",
                     "Fourth quintile"="4",
                     "Fifth quintile"="5",
)

d$Type <- sapply(d$Type, switch, 
                      "Food and non-alcoholic beverages"="FOOD",
                      "Alcoholic beverages, tobacco and narcotics"="ALCOHOL, TOBACCO",
                      "Clothing and footwear"="CLOTHING",
                      "Housing, water, electricity, gas and other fuels"="ELECTRICITY, GAS",
                      "Furnishings, household equipment and routine household maintenance"="FURNISHINGS",
                      "Health"="HEALTH",
                      "Transport"="TRANSPORT",
                      "Communications"="COMMUNICATION",
                      "Recreation and culture" ="RECREATION",
                      "Education"="EDUCATION",
                      "Restaurants and hotels"="RESTAURANTS",
                      "Miscellaneous goods and services"="MISCELLANEOUS",
)
```

Graph plot for consumption share per purpose by quintile.
```{r}
png("cons_per_quintile.png")
p1<-ggplot(d, aes(fill=Type, y=Value, x=Quintile)) + 
  geom_bar(position="stack", stat="identity")+
  facet_grid(~ Date)+
  theme_bw()+
  labs(y="Share","Consumption share over time per quintile", 
       title="Consumption share EA by quintile and year",
       caption = "Author's calculation, Eurostat consumption expenditure by income quintile and COICOP")+ 
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Paired")
dev.off()
p1
```
Compute the average consumption expenditure household weighted price index.

```{r}
colnames(log_annual)<-c("Date","Type","value","Valueindex","y","log_v")

pricequantity <- log_annual %>% right_join(average_con, by=c("Type","y"))

pq<-pricequantity %>%
  group_by(y) %>%
  mutate(exposure=sum(Valueindex*Value, na.rm=TRUE))

pq$Date.x <- gsub('-01-01', '', pq$Date.x, fixed=TRUE)

pq<-pq %>%
    group_by(Type)%>%
    mutate(exposure_average = exposure/exposure[Date.x=="2000"])

pq$Date.x<-as.Date(pq$Date.x, "%Y")
pq<-pq %>% filter(between(Date.x, as.Date('2000-01-01'), as.Date('2020-12-31')))

png("Annual_PIndex.png")
p2<-ggplot(pq, aes(x=Date.x,y=exposure_average, col=Type)) + 
  geom_path()+
  labs(x="Date",y="Price index", title="Annual price index for average household", caption="ECB SDW and Eurostat COICOP, Author's calculation")+
  theme_bw()
dev.off()
  p2
```
Compute the consumption expenditure per purpose per quintile weighted price index.
```{r}
quintile_exposure <- d[order(d$Quintile),]                                            
quintile_exposure$Date<-as.Date(quintile_exposure$Date, format='%Y')
quintile_exposure$Date <- sub('-09-03', '', quintile_exposure$Date, fixed=TRUE)
quintile_exposure$Date <- sub('1999', '2000', quintile_exposure$Date, fixed=TRUE)


log_annual$Date<-as.character(log_annual$Date)
log_annual$Date <- gsub('-01-01', '', log_annual$Date, fixed=TRUE)

total_quintile<-log_annual %>% right_join(quintile_exposure, by=c("Date","Type"))

pq_quintile<-total_quintile %>%
  group_by(Date, income_grp) %>%
  mutate(exposure=sum(Valueindex*Value, na.rm=TRUE))

pq_quintile<-pq_quintile %>%
    group_by(income_grp)%>%
    mutate(exposure_average = exposure/exposure[Date=="2000"])

pq_quintile <- pq_quintile[order(pq_quintile$Quintile),]                                            

png("Pindex_quintile.png")
ggplot(pq_quintile, aes(x=as.Date(Date, "%Y"),y=exposure_average, col=Quintile)) + 
  geom_line()+
  labs(x="Date",y="Price index", title="5year price index by quintile", caption="ECB SDW and Eurostat COICOP, Author's calculation")+
  theme_bw()
dev.off()
p3
```

City type consumption heteroegeneity
```{r}
colnames(cons_per_city)<-c("Type","city_type","s_2005","s_2010","s_2015")

d<-cons_per_city %>%
    group_by(city_type) %>%
    mutate(Somme_2005=sum(s_2005), Somme_2010=sum(s_2010), Somme_2015=sum(s_2015))

share<-d %>%
  group_by(city_type) %>%
  summarize(Type=Type, s_2005/Somme_2005, s_2010/Somme_2010, s_2015/Somme_2015 )

colnames(share)<-c("city_type","Type","s_2005","s_2010","s_2015")
d <- melt(share, id.vars=c("Type","city_type"))
d$variable <- gsub('s_', '', d$variable, fixed=TRUE)

colnames(d)<-c("Type", "city_type", "Date", "Value")

d$Type <- sapply(d$Type, switch, 
                      "Food and non-alcoholic beverages"="FOOD",
                      "Alcoholic beverages, tobacco and narcotics"="ALCOHOL, TOBACCO",
                      "Clothing and footwear"="CLOTHING",
                      "Housing, water, electricity, gas and other fuels"="ELECTRICITY, GAS",
                      "Furnishings, household equipment and routine household maintenance"="FURNISHINGS",
                      "Health"="HEALTH",
                      "Transport"="TRANSPORT",
                      "Communications"="COMMUNICATION",
                      "Recreation and culture" ="RECREATION",
                      "Education"="EDUCATION",
                      "Restaurants and hotels"="RESTAURANTS",
                      "Miscellaneous goods and services"="MISCELLANEOUS",)

d$city_type <- sapply(d$city_type, switch, 
                      "Cities"="City",
                      "Rural areas"="Rural",
                      "Towns and suburbs"="Suburbs",)           
#Plots
png("cons_per_city.png")
ggplot(d, aes(fill=Type, y=Value, x=city_type)) + 
  geom_bar(position="stack", stat="identity")+
  facet_grid(~ Date)+
  theme_bw()+
  labs(y="Share",x="Urbanisation type", 
       title="Consumption share EA by urbanisation degree and year",
       caption = "Author's calculation, Eurostat consumption expenditure by income city type and COICOP",
       fill="Type")+ 
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Paired")
dev.off()
p4
```

```{r}
log_annual$Date<-as.character(log_annual$Date)

log_annual$Date <- sub('-01-01', '', log_annual$Date, fixed=TRUE)

total_city<-log_annual %>% right_join(d, by=c("Date","Type"))

pq_city<-total_city %>%
  group_by(Date, city_type) %>%
  mutate(exposure=sum(Valueindex*Value, na.rm=TRUE))

pq_city<-pq_city %>%
    group_by(city_type)%>%
    mutate(exposure_average = exposure/exposure[Date=="2005"])

pq_city <- pq_city[order(pq_city$city_type),]                                            

png("Pindex_city.png")
ggplot(pq_city, aes(x=as.Date(Date, "%Y"), y=exposure_average, col=city_type)) + 
  geom_line()+
  labs(x="Date",y="Price index", title="5year price index by degree of urbanisation", 
       col='Urbanisation',
       caption="ECB SDW and Eurostat COICOP, Author's calculation"
       )+
  theme_bw()
dev.off()
```

Cons per purpose by csp
```{r}
colnames(cons_per_csp)<-c("Type","CSP_WORK","s_1999","s_2005","s_2010","s_2015", "s_2020")
d<-cons_per_csp %>%
    group_by(CSP_WORK) %>%
    mutate(Somme_1999 = sum(s_1999), Somme_2005=sum(s_2005), Somme_2010=sum(s_2010), Somme_2015=sum(s_2015), Somme_2020=sum(s_2020))

share<-d %>%
  group_by(CSP_WORK) %>%
  summarize(Type=Type, s_1999/Somme_1999, s_2005/Somme_2005, s_2010/Somme_2010, s_2015/Somme_2015, s_2020/Somme_2020)

colnames(share)<-c("CSP_WORK","Type","s_1999","s_2005","s_2010","s_2015", "s_2020")
d <- melt(share, id.vars=c("Type","CSP_WORK"))
d$variable <- gsub('s_', '', d$variable, fixed=TRUE)

colnames(d)<-c("Type", "CSP_WORK", "Date", "Value")

d$Type <- sapply(d$Type, switch, 
                      "Food and non-alcoholic beverages"="FOOD",
                      "Alcoholic beverages, tobacco and narcotics"="ALCOHOL, TOBACCO",
                      "Clothing and footwear"="CLOTHING",
                      "Housing, water, electricity, gas and other fuels"="ELECTRICITY, GAS",
                      "Furnishings, household equipment and routine household maintenance"="FURNISHINGS",
                      "Health"="HEALTH",
                      "Transport"="TRANSPORT",
                      "Communications"="COMMUNICATION",
                      "Recreation and culture" ="RECREATION",
                      "Education"="EDUCATION",
                      "Restaurants and hotels"="RESTAURANTS",
                      "Miscellaneous goods and services"="MISCELLANEOUS",)

d$CSP_WORK <- sapply(d$CSP_WORK, switch, 
                      "Employed persons except employees"="E",
                      "Manual workers in industry and services"="M",
                      "Non-manual workers in industry and services"="N",
                      "Other persons outside the labour force (former name: inactive persons)"="I",
                      "Retired persons"="R",
                      "Unemployed persons"="U",)
                 
#Plots
png("cons_per_csp.png")
ggplot(d, aes(fill=Type, y=Value, x=CSP_WORK)) + 
  geom_bar(position="stack", stat="identity")+
  facet_grid(~ Date)+
  theme_bw()+
  labs(y="Share", x="Activity status", 
       title="Consumption share France by work and year",
       caption = "Author's calculation, Eurostat consumption expenditure by income city type and COICOP")+ 
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Paired")
dev.off()
p6
```
Price index by CSP
```{r}
log_annual$Date<-as.character(log_annual$Date)

log_annual$Date <- sub('-01-01', '', log_annual$Date, fixed=TRUE)
d$Date <- sub('1999', '2000', d$Date, fixed=TRUE)

colnames(log_annual)<-c("Date","Type","value","Valueindex","y","log_v")
total_csp<-log_annual %>% right_join(d, by=c("Date","Type"))

pq_csp<-total_csp %>%
  group_by(Date, CSP_WORK) %>%
  mutate(exposure=sum(Valueindex*Value, na.rm=TRUE))

pq_csp<-pq_csp %>%
    group_by(CSP_WORK)%>%
    mutate(exposure_average = exposure/exposure[Date=="2000"])

pq_csp <- pq_csp[order(pq_csp$CSP_WORK),]                                            

png("Pindex_csp.png")
ggplot(pq_csp, aes(x=as.Date(Date, "%Y"), y=exposure_average, col=CSP_WORK)) + 
  geom_line()+
  labs(x="Date",y="Price index", title="5year price index by activity status in France", caption="ECB SDW and Eurostat COICOP, Author's calculation", col="Activity Status")+
  theme_bw()
dev.off()
```
CONS SHARE PER AGE
```{r}
colnames(cons_per_age)<-c("Type","AGE_RANGE","s_1999","s_2005","s_2010","s_2015")
d<-cons_per_age %>%
    group_by(AGE_RANGE) %>%
    mutate(Somme_1999 = sum(s_1999), Somme_2005=sum(s_2005), Somme_2010=sum(s_2010), Somme_2015=sum(s_2015))

share<-d %>%
  group_by(AGE_RANGE) %>%
  summarize(Type=Type, s_1999/Somme_1999, s_2005/Somme_2005, s_2010/Somme_2010, s_2015/Somme_2015)

colnames(share)<-c("AGE_RANGE","Type","s_1999","s_2005","s_2010","s_2015")
d <- melt(share, id.vars=c("Type","AGE_RANGE"))
d$variable <- gsub('s_', '', d$variable, fixed=TRUE)

colnames(d)<-c("Type", "AGE_RANGE", "Date", "Value")

d$Type <- sapply(d$Type, switch, 
                      "Food and non-alcoholic beverages"="FOOD",
                      "Alcoholic beverages, tobacco and narcotics"="ALCOHOL, TOBACCO",
                      "Clothing and footwear"="CLOTHING",
                      "Housing, water, electricity, gas and other fuels"="ELECTRICITY, GAS",
                      "Furnishings, household equipment and routine household maintenance"="FURNISHINGS",
                      "Health"="HEALTH",
                      "Transport"="TRANSPORT",
                      "Communications"="COMMUNICATION",
                      "Recreation and culture" ="RECREATION",
                      "Education"="EDUCATION",
                      "Restaurants and hotels"="RESTAURANTS",
                      "Miscellaneous goods and services"="MISCELLANEOUS",)

d$AGE_RANGE <- sapply(d$AGE_RANGE, switch, 
                      "Less than 30 years"= "30",
                      "From 30 to 44 years"="30-45",
                      "From 45 to 59 years"="45-59",
                      "60 years or over"="60",)
                 
d <- d[order(d$AGE_RANGE),]                                            

#Plots
png("cons_per_age.png")
ggplot(d, aes(fill=Type, y=Value, x=AGE_RANGE)) + 
  geom_bar(position="stack", stat="identity")+
  facet_grid(~ Date)+
  theme_bw()+
  labs(y="Share",x="Age range", 
       title="Consumption share France by age range and year",
       caption = "Author's calculation, Eurostat consumption expenditure by income city type and COICOP")+ 
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Paired")
dev.off()
```
Price index for age
```{r}
log_annual$Date<-as.character(log_annual$Date)

log_annual$Date <- sub('-01-01', '', log_annual$Date, fixed=TRUE)
d$Date <- sub('1999', '2000', d$Date, fixed=TRUE)

colnames(log_annual)<-c("Date","Type","value","Valueindex","y","log_v")
total_age<-log_annual %>% right_join(d, by=c("Date","Type"))

pq_age<-total_age %>%
  group_by(Date, AGE_RANGE) %>%
  mutate(exposure=sum(Valueindex*Value, na.rm=TRUE))

pq_age<-pq_age %>%
    group_by(AGE_RANGE)%>%
    mutate(exposure_average = exposure/exposure[Date=="2000"])

pq_age <- pq_age[order(pq_age$AGE_RANGE),]                                            

png("Pindex_age.png")
ggplot(pq_age, aes(x=as.Date(Date, "%Y"), y=exposure_average, col=AGE_RANGE)) + 
  geom_line()+
  labs(x="Date",y="Price index", title="5year price index by age in France", caption="ECB SDW and Eurostat COICOP, Author's calculation", col= "Age range")+
  theme_bw()
dev.off()
```
